On the properties of projectile motion and quadratic air resistance

### 2016-04-18 Paper/main.tex
Final

	\section{Introduction}
	Projectile motion is the motion in which an object travels under the constant effect of a force field. The simplest form of projectile motion considers a closed system with no air resistance while a more advanced one might consider multiple gravitational pulls or air resistance. This paper will first describe the ideal trajectory and properties of a projectile parabola and then develop and test an analytical solution to quadratic air resistance.
	\section{Trajectory}
	Consider a Euclidean space $\mathbb{R}^3$ with the three axes $x \comma y$ and $z$. At any time $t$ we want to be able to describe the acceleration, velocity and position of an object with the three vectors $A = (a_x \comma a_y \comma a_z)\comma V = (v_x \comma v_y \comma v_z)$ and $P = (x \comma y \comma z)$. Let $\theta$ be the angle between the $xz$ plane and the initial direction of the object and $\alpha$ the angle between the $x$ axis and a projection of the path on the $xz$ plane. This is visualized in the graph below.
	\begin{center}
		\trimbox{0cm 0cm 0cm -0.6cm} {
			\tdplotsetmaincoords{40}{120} 
			\begin{tikzpicture} [scale = 4, tdplot_main_coords, axis/.style={ -> , black, semithick}, 
			vector/.style={-stealth, black, ultra thick}, 
			vector guide/.style={dashed, black, ultra thick}]

@@

	Here $C$ is a constant of integration and can be written as $v_0 \sin\theta$ due to $\theta$ being the angle between the $xz$ plane and the $y$ axis. This allows us to define the total velocity vector $V$.

@@

	Which is a second degree polynomial proving that projectile motion can be described as a parabola.

@@

	%kanske tydliggöra varför x, z måste separeras och allt
	Where $t_{max}$ is the time the object spends in the air. We can now substitute $t$ for $t_{max}$ in our expression for the $x$ and $z$ positions.

@@

	Let $R$ be the range. It is then true that

@@

	In order to determine the maximum $R$ we will study the derivative of $R$ with respect to $\theta$.

@@

	Solving $\frac{\partial d}{\partial \theta} = 0$ gives the maxima of $R$.

@@

	%Mindre fumlig motivering
	For the purpose of this section we will not consider the $z$ axis nor the angle $\alpha$ between the $x$ and $z$ axis. The reasoning behind this is that is would obstruct the results we're trying to conclude. Know that these results are applicable to three dimensions with minimal adjustments including but not limited to multiplying uses of $x$ with $\cos\alpha$.

@@

	From here we can define the functions $\theta(R, v_0)$ and $v_0(R, \theta)$.

@@

		\theta(R, v_0) &= \frac12\arcsin\Big(\frac{Rg}{v_{0}^2}\Big)\\\nonumber
		v_0(R, \theta) &= \sqrt{\frac{Rg}{\sin2\theta}}

@@

	This can be solved using the quadratic formula which is as follows
	\begin{align}\label{eq:quadFormula}
		ax^2+bx+c=0 \implies x=\frac{-b\pm\sqrt{b^2-4ac}}{2a}
	Substituting $x$ for $h$, $a$ for $\frac{-gx^2}{2v_{0}^2}$, $b$ for $x$ and $c$ for $-\frac{gx^2}{2v_{0}^2}-y$ gives

@@

	Consider a parabola where $y(0) > 0$. We will denote this initial location on the $y$ axis $y_0$. We now want to determine which angle $\theta$ yields the farthest possible parabola. We begin by solving $y=0$ for $t$ where our constant of integration $c_y$ for $y$ is $y_0$.

@@

		y=v_0\sin\theta\,t - \frac12gt^2 + y_0 &= 0 \implies\\\nonumber
		-\frac12g\,t^2+v_0\sin\theta\,t+y_0 &= 0

@@

	Recall equation \eqref{eq:quadFormula}. Substituting $a$ for $-\frac12g$, $b$ for $v_0\sin\theta$ and $c$ for $y_0$ gives the following expression for $t_{max}$

@@

	Setting the second term as negative yields a negative time when $y_0 > 0$ which is nonsense. We will therefore write it as positive from now on. We can now find the length of a parabola and see how it varies with a changing angle $\theta$. Let $R$ denote the range of the projectile and let it be given by $x(t_{max})$.

@@

@@

	We want to write an algorithm that takes the initial height and velocity and returns the optimal angle. The code tests all possible angles $\theta$ in our equation for $R$ and determines which gives the farthest possible throw. The following pseudo code details the procedure.

@@

		\end{algorithmic}
	\end{algorithm}
	The following two plots are the results of the above algorithm with the initial conditions $v_0 = 10\,m/s$, $h_{min} = 0\,m, h_{max} = 20\,m$ and $v_0 = 25\,m/s$, $h_{min} = 0\,m, h_{max} = 100\,m$ respectively.
	\begin{center}
		\begin{tikzpicture}
		\begin{axis}[ 
		xlabel=Starting height (m),
		ylabel=Optimal angle (deg)
		] 
		\addplot[solid] coordinates {(0, 44.9944) (1, 39.9581) (2, 34.9963) (3, 30.0631) (4, 25.1758) (5, 20.4775) (6, 16.3121) (7, 13.0176) (8, 10.6398) (9, 8.9725) (10, 7.7865) (11, 6.9156) (12, 6.2510) (13, 5.7238) (14, 5.2999) (15, 4.9504) (16, 4.6581) (17, 4.4118) (18, 4.1941) (19, 4.0050) (20, 3.8388) };
		\end{axis}
		\end{tikzpicture}
	\end{center}
	\begin{center}
		\begin{tikzpicture}
		\begin{axis}[ 
		xlabel=Starting height (m),
		ylabel=Optimal angle (deg)
		] 
		\addplot[solid] coordinates {(0, 44.9944) (5, 40.6857) (10, 36.3370) (15, 31.8736) (20, 27.2098) (25, 22.2594) (30, 17.0455) (35, 12.0493) (40, 8.3308) (45, 6.1306) (50, 4.8530) (55, 4.0508) (60, 3.5065) (65, 3.1112) (70, 2.8132) (75, 2.5783) (80, 2.3892) (85, 2.2345) (90, 2.1028) (95, 1.9882) (100, 1.8908) };
		\end{axis}
		\end{tikzpicture}
	\end{center}
	These graph both follow the same pattern that the distance gained from aiming upwards shrinks as the starting height increases. This is because the travel time cannot be changed all that much when the starting height already contributes so much to the distance.
	\\~\\
	The full code is available at GitHub\cite{code} and mostly contained in \texttt{math.js}. Algorithm \ref{alg:optimalAngle} can be found in the function \texttt{bestAngleFromStartingHeight()}.
	\section{Computational approach to air resistance}\label{seq:numericalAirRes}
	\subsection{Change at every time interval}
	At every time interval $\Delta\,t$ we can apply a force on our object in an angle opposite to $\beta$ proportional to our velocity $v$. While this does not give us a equation to work with we can use Euler's method to get a numerical approximation.

	Let $F_D$ be the force in the opposite direction of that which our object is traveling. At a time interval $\Delta t$ we will add a drag velocity $v_{drag}$ to our total velocity vector $v$.
	\begin{align}\nonumber
		\vec{v}_{new} = \vec{v}_{old} + \vec{v}_{drag}
	\end{align}
	In order to extrapolate $\vec{v}_{drag}$ from $F_D$ we need to consider our objects mass which we will denote $m$. This is due to 
	\begin{align}\nonumber
		\vec{F}=m\vec{a} = m\frac{\Delta\,\vec{v}}{\Delta\,t}\implies \Delta \vec{v} = \frac{\vec{F} \cdot \Delta\,t}{m} = \vec{v}_{drag}
	\end{align}
	This means is that the air resistance is inversely proportional the object's mass. While this might seem counterintuitive at first it is important to remember that a system with a higher mass means that the total energy of the system goes up. It is therefore only logical that our object travels further with higher mass; there is nowhere else for the extra energy to go. One can also consider conservation of momentum in terms of the air mass.
	\\~\\
	$\vec{F}_D$ is given by the quadratic drag equation
	\begin{align}\nonumber
		\vec{F}_D &= -\frac12\rho\,A\,C_D\,|v|^2\hat{v} \implies\\\nonumber
		\vec{v}_{drag} &= \begin{bmatrix}
			-0.5\,\rho\,A\,C_D\,v\,v_x\,m^{-1}\,\Delta\,t\\
			-0.5\,\rho\,A\,C_D\,v\,v_y\,m^{-1}\,\Delta\,t
		\end{bmatrix}
	\end{align}
	where $\rho$ is the air density, $A$ is the frontal area of our projectile, $m$ is the mass of the object and $C_D$ is the coefficient of drag which depends on the shape of our projectile. As $\Delta \rightarrow 0$ our calculations will be more computationally intensive but have greater accuracy.
	\subsection{Algorithm}
	We want a way to numerically apply a force to an object at every time interval and keep track of its velocity and position. The algorithm used can be summarized as the following.
	\begin{algorithm}[H]
		\caption{Algorithm for plotting projectile parabola with air resistance}
		\label{alg:airResistance}
		\begin{algorithmic}
			\Procedure{Air resistance array}{$v_0 ,\theta, \rho, C_D, m, y_0$}
			\State $g \gets $ $9.82$
			\State $x \gets 0$
			\State $y \gets y_0$
			\State $v_x \gets \cos\theta \cdot v_0$
			\State $v_y \gets \sin\theta \cdot v_0$
			\State $\Delta \, t \gets 0.001$;
			\State result $\gets$ [ ]
			\State
			\While {($y > 0$)}
			\State $v \gets \sqrt{v_x^2 + v_y^2}$
			\State $v_x \gets v_x - 0.5\cdot\rho\cdot C_D\cdot v\cdot v_x \cdot m^{-1}\cdot\Delta\,t$
			\State $v_y \gets v_y - 0.5\cdot\rho\cdot C_D\cdot v\cdot v_y \cdot m^{-1}\cdot\Delta\,t$
			\State
			\State $v_y \gets v_y -g \cdot \Delta \, t$
			\State
			\State $x \gets x + v_x \cdot \Delta\,t$
			\State $y \gets y + v_y \cdot \Delta\,t$
			\State
			\State result $\gets$ result $+$ ($x$, $y$)
			\EndWhile
			\State
			\State return result
			\EndProcedure

@@

	The full code is available at GitHub\cite{code} and mostly contained in \texttt{math.js}. Algorithm \ref{alg:airResistance} can be found in the functions \texttt{getAirResistanceData()} and \texttt{getAirResistance()}.
	\section{Analytical solution to one-dimensional air resistance}\label{seq:oneDim}
	\subsection{Derivation}\label{seq:oneDimIntro}
	Consider the simplest case of air resistance; the one-dimensional falling object. Our object does not travel in a parabola and we have previously described this motion as $y(t) = -\frac12gt^2+y_0$. Knowing the air resistance at any time $t$ we can construct a differential equation describing the motion of our object which is as follows

@@

	For simplicity; let $H = \int \tanh(\sqrt{gk}\,t)\,dt, c = \sqrt{gk}, s = c\,t \implies \frac{ds}{dt} = c \implies dt = \frac{ds}{c}$. The integral is now trivial and we can substitute back $s$ after evaluating it.

@@

		\addplot[solid, dashed] coordinates {(0.000, 4.000) (0.010, 4.000) (0.020, 3.998) (0.030, 3.996) (0.040, 3.992) (0.050, 3.988) (0.060, 3.982) (0.070, 3.976) (0.080, 3.969) (0.090, 3.960) (0.100, 3.951) (0.110, 3.941) (0.120, 3.929) (0.130, 3.917) (0.140, 3.904) (0.150, 3.890) (0.160, 3.874) (0.170, 3.858) (0.180, 3.841) (0.190, 3.823) (0.200, 3.804) (0.210, 3.783) (0.220, 3.762) (0.230, 3.740) (0.240, 3.717) (0.250, 3.693) (0.260, 3.668) (0.270, 3.642) (0.280, 3.615) (0.290, 3.587) (0.300, 3.558) (0.310, 3.528) (0.320, 3.497) (0.330, 3.465) (0.340, 3.432) (0.350, 3.399) (0.360, 3.364) (0.370, 3.328) (0.380, 3.291) (0.390, 3.253) (0.400, 3.214) (0.410, 3.175) (0.420, 3.134) (0.430, 3.092) (0.440, 3.049) (0.450, 3.006) (0.460, 2.961) (0.470, 2.915) (0.480, 2.869) (0.490, 2.821) (0.500, 2.772) (0.510, 2.723) (0.520, 2.672) (0.530, 2.621) (0.540, 2.568) (0.550, 2.515) (0.560, 2.460) (0.570, 2.405) (0.580, 2.348) (0.590, 2.291) (0.600, 2.232) (0.610, 2.173) (0.620, 2.113) (0.630, 2.051) (0.640, 1.989) (0.650, 1.926) (0.660, 1.861) (0.670, 1.796) (0.680, 1.730) (0.690, 1.662) (0.700, 1.594) (0.710, 1.525) (0.720, 1.455) (0.730, 1.383) (0.740, 1.311) (0.750, 1.238) (0.760, 1.164) (0.770, 1.089) (0.780, 1.013) (0.790, 0.936) (0.800, 0.858) (0.810, 0.779) (0.820, 0.699) (0.830, 0.618) (0.840, 0.536) (0.850, 0.453) (0.860, 0.369) (0.870, 0.284) (0.880, 0.198) (0.890, 0.111) (0.900, 0.023) (0.910, -0.066) };
		\end{axis}
		\end{tikzpicture}
	\end{center}
	The slow start with a negative second derivative matches our intuition and at roughly $t = 0.8\,s$ the second derivative appears to approach zero. This should be due to the object reaching its terminal velocity which we can calculate using equation \eqref{eq:oneDimAirResTerminalVel}.
	\begin{align}\nonumber
		v_{max} = \sqrt{\frac{g}{k}} = \sqrt{\frac{9.82}{0.432}} = 4.77\,m/s
	\end{align}
	We can approximate the derivative at $t=0.8$ as $\frac{\Delta y}{\Delta t} \approx 4.4 \approx v_{max}$. Equation \eqref{eq:oneDimAirResTMax} also predicts a $t_{max}$ which can be tested.
	\begin{align}\nonumber
		t_{max} = \frac{\cosh^{-1}(\exp(y_0\,k))}{\sqrt{gk}} = \frac{\cosh^{-1}(\exp(1.728))}{\sqrt{4.424224}} \approx 1.17\,s
	\end{align}
	The plot reports $t_{max} \approx 1.17\,s$. Comparing this to a vacuum solution we can use equation \eqref{eq:tMaxFromHeight} which gives
	\begin{align}\nonumber
		t_{max} &= \frac{v_0\sin\theta}{g}+\frac{\sqrt{v_{0}^2\sin^2\theta+2g\,y_0}}{g}\\\nonumber
		&= \frac{\sqrt{2\,g\,y_0}}{g} = \frac{\sqrt{78.56}}{9.82} \approx 0.90\,s
	\end{align}
	This is a major discrepancy which adds credit to the accuracy of equation \eqref{eq:oneDimAirResTMax}.

	We can plot another $y(t)$ where $k$ is much smaller which should result in our object not reaching its terminal velocity in such short time. The following measurements are from a small steel ball with $k = 0.00313$.
	\begin{align}\nonumber
		\begin{cases}\nonumber
			C_D &= 0.47 \\
			m &= 0.022\,kg\\
			y_0 &= 3\,m\\
			v_0 &= 0\,m/s\\
			\rho &= 1.225\,kg/m^3\\
			A &= 0.00023916\,m^2
		\end{cases} \implies k = 0.00313
	\end{align}
	\begin{center}
		\begin{tikzpicture}
		\begin{axis}[ xlabel=t (s), ylabel=y (m)]
		\addplot[solid] coordinates {(0.000, 3.000) (0.010, 3.000) (0.020, 2.998) (0.030, 2.996) (0.040, 2.992) (0.050, 2.988) (0.060, 2.982) (0.070, 2.976) (0.080, 2.969) (0.090, 2.960) (0.100, 2.951) (0.110, 2.941) (0.120, 2.929) (0.130, 2.917) (0.140, 2.904) (0.150, 2.890) (0.160, 2.874) (0.170, 2.858) (0.180, 2.841) (0.190, 2.823) (0.200, 2.804) (0.210, 2.784) (0.220, 2.762) (0.230, 2.740) (0.240, 2.717) (0.250, 2.693) (0.260, 2.668) (0.270, 2.642) (0.280, 2.615) (0.290, 2.587) (0.300, 2.558) (0.310, 2.528) (0.320, 2.497) (0.330, 2.466) (0.340, 2.433) (0.350, 2.399) (0.360, 2.364) (0.370, 2.328) (0.380, 2.292) (0.390, 2.254) (0.400, 2.215) (0.410, 2.175) (0.420, 2.135) (0.430, 2.093) (0.440, 2.050) (0.450, 2.007) (0.460, 1.962) (0.470, 1.917) (0.480, 1.870) (0.490, 1.823) (0.500, 1.774) (0.510, 1.725) (0.520, 1.674) (0.530, 1.623) (0.540, 1.570) (0.550, 1.517) (0.560, 1.463) (0.570, 1.407) (0.580, 1.351) (0.590, 1.294) (0.600, 1.236) (0.610, 1.176) (0.620, 1.116) (0.630, 1.055) (0.640, 0.993) (0.650, 0.930) (0.660, 0.866) (0.670, 0.801) (0.680, 0.735) (0.690, 0.668) (0.700, 0.600) (0.710, 0.531) (0.720, 0.461) (0.730, 0.391) (0.740, 0.319) (0.750, 0.246) (0.760, 0.172) (0.770, 0.098) (0.780, 0.022) (0.790, -0.055) };

@@

			u			&if\,\,\, u \gg w\\
			w			&if\,\,\, w \gg w\\
			\sqrt{2}u	&if\,\,\, u \approx w\\

@@

	We treat these as three separate cases and label them as LAT, HAT and SAT for Low Angle Trajectory, High Angle Trajectory and Split Angle Trajectory respectively.

@@

	We can now plus $u$ from equation \eqref{eq:twoDimU} into equation \eqref{eq:twoDimEMOLATW} and solve for $w$.

@@

	For $y$ the procedure is essentially the same.

>>>>

	\section{Introduction}
	Projectile motion is the motion in which an object travels under the constant effect of a force field. The simplest form of projectile motion considers a closed system with no air resistance while a more advanced one might consider multiple gravitational pulls or air resistance. This paper will first describe the ideal trajectory and properties of a projectile parabola and then develop and test an analytical solution to quadratic air resistance.
	\section{Trajectory}
	Consider a Euclidean space $\mathbb{R}^3$ with the three axes $x \comma y$ and $z$. At any time $t$ we want to be able to describe the acceleration, velocity and position of an object with the three vectors $A = (a_x \comma a_y \comma a_z)\comma V = (v_x \comma v_y \comma v_z)$ and $P = (x \comma y \comma z)$. Let $\theta$ be the angle between the $xz$ plane and the initial direction of the object and $\alpha$ the angle between the $x$ axis and a projection of the path on the $xz$ plane. This is visualized in the figure below.
	\begin{center}
		\trimbox{0cm 0cm 0cm -0.6cm} {
			\tdplotsetmaincoords{40}{120} 
			\begin{tikzpicture} [scale = 4, tdplot_main_coords, axis/.style={ -> , black, semithick}, 
			vector/.style={-stealth, black, ultra thick}, 
			vector guide/.style={dashed, black, ultra thick}]

@@

	Here $C$ is a constant of integration and can be written as $v_0 \sin\theta$ due to $\theta$ being the angle between the $xz$ plane and the $y$ axis making it the initial velocity in the $y$ direction. This allows us to define the total velocity vector $V$.

@@

	Which is a second degree polynomial.

@@

	Where $t_{max}$ is the time the object spends in the air. The solution $t=0$ can be discarded as it is the initial value. We can now substitute $t$ for $t_{max}$ in our expression for the $x$ and $z$ positions.

@@

	Let $R$ be the range. By the Pythagorean theorem it is then true that

@@

	The the optimal theta for maximizing R may be determined by studying the derivative of R with respect to theta.

@@

	Solving $\frac{\partial R}{\partial \theta} = 0$ gives the maxima of $R$.

@@

	For the purpose of this section we will not consider the $z$ axis nor the angle $\alpha$ between the $x$ and $z$ axis. The reasoning behind this is that is would introduce unnecessary complexity to the results. Know that these results may be expanded to three dimensions with minimal adjustments for example by multiplying uses of $x$ with $\cos\alpha$.

@@

	From here we can develop expressions for $\theta$ and $v_0$.

@@

		\theta &= \frac12\arcsin\Big(\frac{Rg}{v_{0}^2}\Big)\\\nonumber
		v_0 &= \sqrt{\frac{Rg}{\sin2\theta}}

@@

	This is a second degree polynomial meaning we can solve it for $h$ using the quadratic formula. Substituting $a$ for $\frac{-gx^2}{2v_{0}^2}$, $b$ for $x$ and $c$ for $-\frac{gx^2}{2v_{0}^2}-y$ gives

@@

	Consider a parabola where $y(0) > 0$. We will denote this initial location on the $y$ axis $y_0$. We now want to determine which angle $\theta$ yields the farthest possible parabola. We begin by solving $y=0$ for $t_{max}$ where our constant of integration $c_y$ for $y$ is $y_0$.

@@

		y=v_0\sin\theta\,t_{max} - \frac12gt_{max}^2 + y_0 &= 0 \implies\\\nonumber
		-\frac12g\,t_{max}^2+v_0\sin\theta\,t_{max}+y_0 &= 0

@@

	Using the quadratic formula and substituting $a$ for $-\frac12g$, $b$ for $v_0\sin\theta$ and $c$ for $y_0$ we obtain the following expression for $t_{max}$

@@

	Setting the second term as negative yields a negative time since $y_0$ is defined as $y_0 > 0$. This is a representation of the object traveling in a parabola at all times $t$ and crossing the initial position at $t=0$. We will therefore write it as positive from now on. We can now find the length of a parabola and see how it varies with a changing angle $\theta$. Let $R$ denote the range of the projectile given by $x(t_{max})$.

@@

	Maximizing $R$ by finding an optimal $\theta$ is not easily solvable algebraically. We will therefore move on to numerical methods. 

@@

	We want to write an algorithm that takes the initial height and velocity and returns the optimal angle. The code tests as many as possible angles $\theta$ in our equation for $R$ and determines which gives the farthest possible throw. The procedure is described in the following pseudo code.

@@

		\end{algorithmic}
	\end{algorithm}
	The following two plots are the results of the above algorithm with the initial conditions $v_0 = 10\,m/s$, $h_{min} = 0\,m, h_{max} = 20\,m$ and $v_0 = 25\,m/s$, $h_{min} = 0\,m, h_{max} = 100\,m$ respectively.
	\begin{center}
		\begin{tikzpicture}
		\begin{axis}[ 
		xlabel=Starting height (m),
		ylabel=Optimal angle (deg)
		] 
		\addplot[solid] coordinates {(0, 44.9944) (1, 39.9581) (2, 34.9963) (3, 30.0631) (4, 25.1758) (5, 20.4775) (6, 16.3121) (7, 13.0176) (8, 10.6398) (9, 8.9725) (10, 7.7865) (11, 6.9156) (12, 6.2510) (13, 5.7238) (14, 5.2999) (15, 4.9504) (16, 4.6581) (17, 4.4118) (18, 4.1941) (19, 4.0050) (20, 3.8388) };
		\end{axis}
		\end{tikzpicture}
	\end{center}
	\begin{center}
		\begin{tikzpicture}
		\begin{axis}[ 
		xlabel=Starting height (m),
		ylabel=Optimal angle (deg)
		] 
		\addplot[solid] coordinates {(0, 44.9944) (5, 40.6857) (10, 36.3370) (15, 31.8736) (20, 27.2098) (25, 22.2594) (30, 17.0455) (35, 12.0493) (40, 8.3308) (45, 6.1306) (50, 4.8530) (55, 4.0508) (60, 3.5065) (65, 3.1112) (70, 2.8132) (75, 2.5783) (80, 2.3892) (85, 2.2345) (90, 2.1028) (95, 1.9882) (100, 1.8908) };
		\end{axis}
		\end{tikzpicture}
	\end{center}
	These graph both follow the same pattern that the distance gained from aiming upwards shrinks as the starting height increases. This is because the travel time cannot be changed all that much when the starting height already contributes so much to the distance.
	\\~\\
	The full code is available at GitHub\cite{code} and mostly contained in \texttt{math.js}. Algorithm \ref{alg:optimalAngle} can be found in the function \texttt{bestAngleFromStartingHeight()}.
	\section{Computational approach to air resistance}\label{seq:numericalAirRes}
	\subsection{Change at every time interval}
	Air resistance is in some way proportional to the velocity of the object and directed in the direction opposite to the current movement. For our algorithm this means that at every time interval $\Delta\,t$ we want to apply a force on our object in an angle opposite to $\beta$ in some way proportional to our velocity $v$.

	Let $F_D$ be the force in the opposite direction of that which our object is traveling. At a time interval $\Delta t$ we will add a drag velocity $v_{drag}$ to our total velocity vector $v$.
	\begin{align}\nonumber
		\vec{v}_{new} = \vec{v}_{old} + \vec{v}_{drag}
	\end{align}
	In order to determine $\vec{v}_{drag}$ from $F_D$ we need to consider our objects mass which we will denote $m$. This is due to 
	\begin{align}\nonumber
		\vec{F}=m\vec{a} = m\frac{\Delta\,\vec{v}}{\Delta\,t}\implies \Delta \vec{v} = \frac{\vec{F} \cdot \Delta\,t}{m} = \vec{v}_{drag}
	\end{align}
	This means is that the air resistance is inversely proportional the object's mass. While this might seem counterintuitive at first it is important to remember that a system with a higher mass means that the total energy of the system increases. It is therefore logical that our object travels further with higher mass; there is nowhere else for the extra energy to go. One can also consider conservation of momentum in terms of the air mass.
	\\~\\
	$\vec{F}_D$ is given by the quadratic drag equation
	\begin{align}\nonumber
		\vec{F}_D &= -\frac12\rho\,A\,C_D\,|v|^2\hat{v} \implies\\\nonumber
		\vec{v}_{drag} &= \begin{bmatrix}
			-0.5\,\rho\,A\,C_D\,v\,v_x\,m^{-1}\,\Delta\,t\\
			-0.5\,\rho\,A\,C_D\,v\,v_y\,m^{-1}\,\Delta\,t
		\end{bmatrix}
	\end{align}
	where $\rho$ is the air density, $A$ is the frontal area of our projectile, $m$ is the mass of the object and $C_D$ is the coefficient of drag which depends on the shape of our projectile. As $\Delta \rightarrow 0$ our calculations will be more computationally intensive but have greater accuracy.
	\subsection{Algorithm}
	We want a way to numerically apply a force to an object at every time interval and keep track of its velocity and position. The algorithm used can be summarized as the following.
	\begin{algorithm}[H]
		\caption{Algorithm for plotting projectile parabola with air resistance}
		\label{alg:airResistance}
		\begin{algorithmic}
			\Procedure{Air resistance array}{$v_0 ,\theta, \rho, C_D, m, y_0$}
			\State $g \gets $ $9.82$
			\State $x \gets 0$
			\State $y \gets y_0$
			\State $v_x \gets \cos\theta \cdot v_0$
			\State $v_y \gets \sin\theta \cdot v_0$
			\State $\Delta \, t \gets 0.001$;
			\State result $\gets$ [ ]
			\State
			\While {($y > 0$)}
			\State $v \gets \sqrt{v_x^2 + v_y^2}$
			\State $v_x \gets v_x - 0.5\cdot\rho\cdot C_D\cdot v\cdot v_x \cdot m^{-1}\cdot\Delta\,t$
			\State $v_y \gets v_y - 0.5\cdot\rho\cdot C_D\cdot v\cdot v_y \cdot m^{-1}\cdot\Delta\,t$
			\State
			\State $v_y \gets v_y -g \cdot \Delta \, t$
			\State
			\State $x \gets x + v_x \cdot \Delta\,t$
			\State $y \gets y + v_y \cdot \Delta\,t$
			\State
			\State result $\gets$ result $+$ ($x$, $y$)
			\EndWhile
			\State
			\State return result
			\EndProcedure

@@

		\end{axis}
		\end{tikzpicture}
	\end{center}
	The full code is available at GitHub\cite{code} and mostly contained in \texttt{math.js}. Algorithm \ref{alg:airResistance} can be found in the functions \texttt{getAirResistanceData()} and \texttt{getAirResistance()}.
	\section{Analytical solution to one-dimensional air resistance}\label{seq:oneDim}
	\subsection{Derivation}\label{seq:oneDimIntro}

	Consider the simplest case of air resistance; the one-dimensional falling object. Our object does not travel in a parabola and we have previously described this motion as $y(t) = -\frac12gt^2+y_0$. Knowing the air resistance at any time $t$ we can construct a differential equation describing the motion of our object

@@

	For simplicity; let $H = \int \tanh(\sqrt{gk}\,t)\,dt \comma c = \sqrt{gk} \comma s = c\,t \implies \frac{ds}{dt} = c \implies dt = \frac{ds}{c}$. The integral is now trivial and we can substitute back $s$ after evaluating it.

@@

		\addplot[solid, dashed] coordinates {(0.000, 4.000) (0.010, 4.000) (0.020, 3.998) (0.030, 3.996) (0.040, 3.992) (0.050, 3.988) (0.060, 3.982) (0.070, 3.976) (0.080, 3.969) (0.090, 3.960) (0.100, 3.951) (0.110, 3.941) (0.120, 3.929) (0.130, 3.917) (0.140, 3.904) (0.150, 3.890) (0.160, 3.874) (0.170, 3.858) (0.180, 3.841) (0.190, 3.823) (0.200, 3.804) (0.210, 3.783) (0.220, 3.762) (0.230, 3.740) (0.240, 3.717) (0.250, 3.693) (0.260, 3.668) (0.270, 3.642) (0.280, 3.615) (0.290, 3.587) (0.300, 3.558) (0.310, 3.528) (0.320, 3.497) (0.330, 3.465) (0.340, 3.432) (0.350, 3.399) (0.360, 3.364) (0.370, 3.328) (0.380, 3.291) (0.390, 3.253) (0.400, 3.214) (0.410, 3.175) (0.420, 3.134) (0.430, 3.092) (0.440, 3.049) (0.450, 3.006) (0.460, 2.961) (0.470, 2.915) (0.480, 2.869) (0.490, 2.821) (0.500, 2.772) (0.510, 2.723) (0.520, 2.672) (0.530, 2.621) (0.540, 2.568) (0.550, 2.515) (0.560, 2.460) (0.570, 2.405) (0.580, 2.348) (0.590, 2.291) (0.600, 2.232) (0.610, 2.173) (0.620, 2.113) (0.630, 2.051) (0.640, 1.989) (0.650, 1.926) (0.660, 1.861) (0.670, 1.796) (0.680, 1.730) (0.690, 1.662) (0.700, 1.594) (0.710, 1.525) (0.720, 1.455) (0.730, 1.383) (0.740, 1.311) (0.750, 1.238) (0.760, 1.164) (0.770, 1.089) (0.780, 1.013) (0.790, 0.936) (0.800, 0.858) (0.810, 0.779) (0.820, 0.699) (0.830, 0.618) (0.840, 0.536) (0.850, 0.453) (0.860, 0.369) (0.870, 0.284) (0.880, 0.198) (0.890, 0.111) (0.900, 0.023) (0.910, -0.066) };
	The slow start with a negative second derivative matches our intuition and at roughly $t = 0.8\,s$ the second derivative appears to approach zero. This should be due to the object reaching its terminal velocity which we can calculate using equation \eqref{eq:oneDimAirResTerminalVel}.
		v_{max} = \sqrt{\frac{g}{k}} = \sqrt{\frac{9.82}{0.432}} = 4.77\,m/s
	We can approximate the derivative at $t=0.8$ as $\frac{\Delta y}{\Delta t} \approx 4.4 \approx v_{max}$. Equation \eqref{eq:oneDimAirResTMax} also predicts a $t_{max}$ which can be tested.
		t_{max} = \frac{\cosh^{-1}(\exp(y_0\,k))}{\sqrt{gk}} = \frac{\cosh^{-1}(\exp(1.728))}{\sqrt{4.424224}} \approx 1.17\,s
	The plot reports $t_{max} \approx 1.17\,s$. Comparing this to a vacuum solution we can use equation \eqref{eq:tMaxFromHeight} which gives
		t_{max} &= \frac{v_0\sin\theta}{g}+\frac{\sqrt{v_{0}^2\sin^2\theta+2g\,y_0}}{g}\\\nonumber
		&= \frac{\sqrt{2\,g\,y_0}}{g} = \frac{\sqrt{78.56}}{9.82} \approx 0.90\,s
	This increase of approximately $30\,\%$ matches our physical intuition for such a large object and adds credit to the accuracy of equation \eqref{eq:oneDimAirResTMax}.
	We can plot another $y(t)$ where $k$ is much smaller which should result in our object not reaching its terminal velocity in such short time. The following measurements are from a small steel ball with $k = 0.00313$.
			m &= 0.022\,kg\\
			y_0 &= 3\,m\\
			A &= 0.00023916\,m^2
		\end{cases} \implies k = 0.00313
		\addplot[solid] coordinates {(0.000, 3.000) (0.010, 3.000) (0.020, 2.998) (0.030, 2.996) (0.040, 2.992) (0.050, 2.988) (0.060, 2.982) (0.070, 2.976) (0.080, 2.969) (0.090, 2.960) (0.100, 2.951) (0.110, 2.941) (0.120, 2.929) (0.130, 2.917) (0.140, 2.904) (0.150, 2.890) (0.160, 2.874) (0.170, 2.858) (0.180, 2.841) (0.190, 2.823) (0.200, 2.804) (0.210, 2.784) (0.220, 2.762) (0.230, 2.740) (0.240, 2.717) (0.250, 2.693) (0.260, 2.668) (0.270, 2.642) (0.280, 2.615) (0.290, 2.587) (0.300, 2.558) (0.310, 2.528) (0.320, 2.497) (0.330, 2.466) (0.340, 2.433) (0.350, 2.399) (0.360, 2.364) (0.370, 2.328) (0.380, 2.292) (0.390, 2.254) (0.400, 2.215) (0.410, 2.175) (0.420, 2.135) (0.430, 2.093) (0.440, 2.050) (0.450, 2.007) (0.460, 1.962) (0.470, 1.917) (0.480, 1.870) (0.490, 1.823) (0.500, 1.774) (0.510, 1.725) (0.520, 1.674) (0.530, 1.623) (0.540, 1.570) (0.550, 1.517) (0.560, 1.463) (0.570, 1.407) (0.580, 1.351) (0.590, 1.294) (0.600, 1.236) (0.610, 1.176) (0.620, 1.116) (0.630, 1.055) (0.640, 0.993) (0.650, 0.930) (0.660, 0.866) (0.670, 0.801) (0.680, 0.735) (0.690, 0.668) (0.700, 0.600) (0.710, 0.531) (0.720, 0.461) (0.730, 0.391) (0.740, 0.319) (0.750, 0.246) (0.760, 0.172) (0.770, 0.098) (0.780, 0.022) (0.790, -0.055) };

@@

			u			&\textrm{if}\,\,\, u \gg w\\
			w			&\textrm{if}\,\,\, w \gg w\\
			\sqrt{2}u	&\textrm{if}\,\,\, u \approx w\\

@@

	We treat these as three separate cases and label them as LAT, HAT and SAT for Low Angle Trajectory, High Angle Trajectory and Split Angle Trajectory respectively. At times none of these approximations may be applicable which is their weakness. Additional cases such as $v \approx \sqrt{5}u \,\, \textrm{if} \,\, u \approx 2w$ could be considered but their derivation would be nearly identical to that of $v \approx \sqrt{5}u$.

@@

	We can now plug $u$ from equation \eqref{eq:twoDimU} into equation \eqref{eq:twoDimEMOLATW} and solve for $w$.

@@

	For $y$ we perform essentially the same procedure.
